---
title: "Assignment 5"
author: "Kaspar Sakmann"
date: "20. Dezember 2015"
output: html_document
---

```{r}
#setwd("~/coursera/datasci_course_materials/assignment5/")
library('caret')
dat<-read.csv("seaflow_21min.csv")
```

Q2: how many labeled synecho
```{r}
summary(dat$pop)
```

Q3: 3rd quantile of fsc_small
```{r}
summary(dat$fsc_small)
```

Q4:
```{r}
set.seed(107)
inTrain<-createDataPartition(y=dat$pop,p=0.5, list=FALSE)
str(inTrain)
training<-dat[inTrain,]
testing<-dat[-inTrain,]
nrow(training)
nrow(testing)
mean(testing$time)
```


Q5: What other two categories are the data points labeled 'ultra' somehow mixed with? 
```{r, echo=FALSE}
library(ggplot2)
ggplot(dat,aes(dat$chl_small,dat$pe,color=dat$pop))+geom_point()
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

Q6: which populations cannot be recognized?
```{r}
library(rpart)
fol<-formula(pop ~ fsc_small + fsc_perp + fsc_big + pe + chl_big + chl_small)
tree_model<-rpart(formula=fol,method="class", data=training)
print(tree_model)
```
crypto does not appear on any leaf.

Q7: What is the value for the threshold of the pe field? 5006.5
Q8: Which variables appear to be most important in predicting the class population? pe and chl_small.


Q9: How much of the testing data gets correctly classified?
```{r}
tree_predictions<-predict(tree_model, testing, type='class')
n_correct<-sum(tree_predictions==testing$pop)
n_total<-nrow(testing)
correct_fraction<-n_correct/n_total
correct_fraction
```

Q10: same thing with a random forest
```{r}
library(randomForest)
forest_model<-randomForest(formula=fol,method="class", data=training)
forest_predictions<-predict(forest_model, testing, type='class')
n_correct<-sum(forest_predictions==testing$pop)
n_total<-nrow(testing)
correct_fraction<-n_correct/n_total
correct_fraction
```

Q11: which variables are most important? largest Gini decrease?
```{r}
importance(forest_model)
```
1. pe, 2. chl_small, 3. chl_big

Q12: same thing with a support vector machine. 
```{r}
library(e1071)
svm_model<-svm(formula=fol,method="class", data=training)
svm_predictions<-predict(svm_model, testing, type='class')
n_correct<-sum(svm_predictions==testing$pop)
n_total<-nrow(testing)
correct_fraction<-n_correct/n_total
correct_fraction
```

Q13: confusion matrices of the three models
```{r}
table(pred=tree_predictions, true=testing$pop)
table(pred=forest_predictions, true=testing$pop)
table(pred=svm_predictions, true=testing$pop)
```
The worst error is on mistaking pico for ultra.

Q14: what is the net change in accuracy if the file with file_id 208 is left out on the svm model. 
```{r}
purgeddat=dat[dat$file_id != 208,]
set.seed(107)
inTrain<-createDataPartition(y=purgeddat$pop,p=0.5, list=FALSE)
str(inTrain)
purgedtraining<-purgeddat[inTrain,]
purgedtesting<-purgeddat[-inTrain,]

purgedsvm_model<-svm(formula=fol,method="class", data=purgedtraining)
purgedsvm_predictions<-predict(purgedsvm_model, purgedtesting, type='class')
purgedn_correct<-sum(purgedsvm_predictions==purgedtesting$pop)
purgedn_total<-nrow(purgedtesting)
purgedcorrect_fraction<-n_correct/n_total
purgedcorrect_fraction

improvement<-purgedcorrect_fraction-correct_fraction
improvement
```

Q15: which variable takes only a few discrete values?
```{r}
plot(dat$fsc_big)
```



